---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# dint

[![CRAN status](https://www.r-pkg.org/badges/version/dint)](https://cran.r-project.org/package=dint)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![Travis build status](https://travis-ci.com/s-fleck/dint.svg?branch=master)](https://travis-ci.org/s-fleck/com)
[![Coverage status](https://codecov.io/gh/s-fleck/dint/branch/master/graph/badge.svg)](https://codecov.io/github/s-fleck/dint?branch=master)

A Toolkit for Year-Quarter, Year-Month and Year-Isoweek Dates

S3 classes and methods to create and work with year-quarter and year-month 
vectors. Basic arithmetic operations (such as adding and subtracting) are
supported, as well as formatting and converting to and from standard R
Date types. For more info please refer to the
[package vignette](https://CRAN.R-project.org/package=dint/vignettes/dint.html)
or the [documentation](https://s-fleck.github.io/dint/)


## Dependencies

dint is implemented strictly in base R and will **always stay free of hard dependencies**.
The optional dependencies of dint are just there to facilitate interoperability 
these packages if you are already using them; for example by providing ggplot2 
scales.


## Installation

Install the release version of dint from CRAN:

```{r cran-installation, eval = FALSE}
install.packages("dint")
```

Or install the development version from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("s-fleck/dint")
```





## Example

```{r example}
library(dint)

# creation
q <- date_yq(2014, 4)
m <- as_date_ym(201412)
w <- as_date_yw(as.Date("2017-01-01"))

# printing
print(q)
print(m)
print(w)  # isoweeks do not follow calendar years!

# arithmetic operations
# quarters
q
q + 1
seq(q -2, q + 2)


# months
m
m + 1
seq(m -2, m + 2)


# formatting
format(q)
format(q, "%Y.%q")
format(q, "%y.%q")
format(m)


# get start and end of period
last_of_quarter(q)
first_of_quarter(q)
first_of_month(Sys.Date())
first_of_isoweek(w)
last_of_isoweek(w)

```
