---
title: "Dint: Integer Representations of Calendar Periods"
author: "Stefan Fleck"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{dint}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 3.4,
  fig.height = 2.4,
  comment = "#>"
)
```

```{r}
library(dint)
```

dint helps you with working with year-quarter, year-month and year-isoweek 
dates. It stores them in an easily human readable integer format, e.q `20141` 
for the first quarter of 2014 and so forth. If you are already using such 
integers to represent dates, dint can make many common operations easier for 
you.

dint is implemented in base R and comes with zero external dependencies. 
Even if you don't work with such special dates directly, dint can still 
help you at formatting dates, labelling plot axes, or getting first / last days 
of calendar periods (quarters, months, isoweeks).




# date_xx Vectors

dint provides 4 different S3 classes that inherit from `date_xx`^[`date_xx` is 
just a superclass for all dint date classes, you do not need to use it directly]. 

* `date_yq` for year-quarter dates
* `date_yw` for year-month dates
* `date_yw` for year-isoweek dates. Please note that the year for isoweeks does 
  not necessarily correspond to the calendar year 
  [wikipedia](https://en.wikipedia.org/wiki/ISO_week_date)
* `date_y` for storing years. This class exists for consistency and provides 
  little advantage over storing years as bare integers. The main use of this 
  class is in package development when you want to write your own generics
  and methods for years. 
  
  
  

# Creation

`date_xx` vectors can be created using explicit constructors... 
```{r}
date_yq(2015, 1)
date_ym(c(2015, 2016), c(1, 2))
date_yw(c(2008, 2009), 1)
```

..or through coercion of `Dates` or `integers`
```{r}
as_date_yq(Sys.Date())
as_date_yq(20141)   # the last digit is interpreted as quarter
as_date_ym(201412)  # the last two digits are interpreted as month
as_date_yw("2018-01-01")  # anything else that can be parsed by as.Date() works
```




# Coercion

You can coerce `Dates` to any `date_xx` subclass with `as_date_**()`
```{r}
d <- as.Date("2018-05-12")
as_date_yq(d)
as_date_ym(d)
as_date_yw(d)
as_date_y(d)
```

Conversely, you can convert `date_xx` back to R `Dates`
```{r}
q <- date_yq(2015, 1)
as.Date(q)
as.POSIXlt(q)
```

`as.POSIXct()` creates datetimes in UTC/GMT, so the result might not always be 
as expected, depending on your local timezone.
```{r}
as.POSIXct(q)
as.POSIXct(q, tz = "GMT")
print(as.POSIXct(q), tz = "GMT")
print(as.POSIXct(q), tz = "CET")

```



# Arithmetic and Sequences

All `date_xx` support addition, subtraction and sequence generation.

```{r}
q <- date_yq(2014, 4)
q + 1
q - 1
seq(q - 2, q + 2)

m <- date_ym(2014, 12)
m + 1
m - 1
seq(m - 2, m + 2)

w <- date_yw(2017, 33)
w + 1
w - 1
seq(w - 2, w + 2)

```




# Accessors

## date_xx Components

You can access components of `date_xx` (e.g the quarter of a `date_yq`) with 
accessor functions. You can also use these functions to convert between
`date_xx` vectors.

```{r}
q <- date_yq(2014, 4)
get_year(q)
get_quarter(q)
get_month(q) # defaults to first month of quarter
get_isoweek(q)

m <- date_ym(2014, 12)
get_year(m)
get_quarter(m)
get_month(m)
get_isoweek(m)

w <- date_yw(2014, 33)
get_year(w)
get_quarter(w)
get_month(w)
get_isoweek(w)
```


## lubridate Compatibility Accessors

If you use lubridate, you can just use the slightly less verbose lubridate 
accessors

```{r}
suppressPackageStartupMessages(library(lubridate))

year(q)
quarter(q)
month(q)
```


You can get the first and last days of calendar periods with dint
```{r}
q <- date_yq(2015, 1)
first_of_quarter(q)  # the same as as.Date(q), but more explicit
last_of_quarter(q)  # the same as as.Date(q), but more explicit
```

These functions work with normal dates
```{r}
d <- as.Date("2018-05-12")
first_of_year(d)
last_of_year(d)
first_of_quarter(d)
last_of_quarter(d)
first_of_month(d)
last_of_month(d)
first_of_isoweek(d)
last_of_isoweek(d)

# Alternativeley you can use these:
first_of_yq(2012, 2)
last_of_ym(2012, 2)
last_of_yw(2011, 52)
```




# Formatting

## format() Methods

Formatting `date_xx` vectors is easy and uses a subset of the placeholders of 
`base::strptime()` (+ `%q` for quarters).

```{r}
q <- date_yq(2014, 4)
format(q, "%Y-Q%q")  # iso/default
format(q, "%Y.%q")
format(q, "%y.%q")

m <- date_ym(2014, 12)
format(m, "%Y-M%m")  # iso/default

w <- date_yw(2014, 1)
format(w, "%Y-W%V")  # iso/default
```


## Shorthands

dint also provides all-in-one functions for common tasks
```{r}
format_yq(Sys.Date())
format_yq_short(Sys.Date())
format_yq_shorter(Sys.Date())

format_ym(Sys.Date())
format_yw(Sys.Date())
```


# ggplot2: Axis Labels

```{r} 
library(ggplot2)
```


## Labelling date_xx Vectors

dint implements `scale_date_**()` and `date_**_breaks()` that provide nicely 
labeled axes for ggplots by default

```{r, collapse=TRUE, fig.show='hold' } 
q <- data.frame(
  time  = seq(date_yq(2016, 1), date_yq(2016, 4)),
  value = rnorm(4)
)

m <- data.frame(
  time  = seq(date_ym(2016, 8), date_ym(2016, 11)),
  value = rnorm(4)
)

w <- data.frame(
  time  = seq(date_yw(2016, 48), date_yw(2017, 1)),
  value = abs(rnorm(6))
)

w2 <- data.frame(
  time  = seq(date_yw(2016, 1), date_yw(2019, 1)),
  value = abs(rnorm(157))
)

ggplot(q, aes(x = time, y = value))  + 
  geom_point()

ggplot(m, aes(x = time, y = value))  + 
  geom_line() +
  scale_x_date_ym(labels = format_ym_shorter)

ggplot(w, aes(x = time, y = value))  + 
  geom_col()  + 
  scale_x_date_yw(labels = format_yw_iso)

ggplot(w2, aes(x = time, y = value)) + 
  geom_smooth() 
```


## Labelling Date Vectors

If you use R `Date` vectors, you can still use the formatting functions supplied
by dint to generate nice axis labels.

```{r, collapse=TRUE, fig.show='hold' } 
x <- data.frame(
  time  = seq(as.Date("2016-01-01"), as.Date("2016-08-08"), by = "day"),
  value = rnorm(221)
)

p <- ggplot(
  x,
  aes(
    x = time, 
    y = value)
  ) + geom_point()

p + ggtitle("iso") + ggtitle("default")
p + scale_x_date(labels = format_yq_iso) + ggtitle("date_yq_iso")
p + scale_x_date(labels = format_ym_short) + ggtitle("date_ym_short")
p + scale_x_date(labels = format_yw_shorter) + ggtitle("date_yw_shorter")
```
